---
title: "Homework 1"
author: Yaa Ababio
output: github_document
---

Loading the tidyverse into R.

```{r load_tidyverse}
library(tidyverse)
```


## Problem 1

#### Creating a data frame
Create a data frame comprised of: a random sample of size 10 from a standard Normal distribution,
a logical vector indicating whether elements of the sample are greater than 0, a character vector of length 10, and a factor vector of length 10 (with 3 different factor “levels”).

```{r data_frame}
prob1_df = 
  tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vec = factor(c("small", "small", "small", "med", "med", "med", "med", "large", "large", "large"))
  )
```




#### Means of variables in the data frame
Take the mean of each variable in the data frame.
```{r mean}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```

Note that R is able to take the means of the numeric and logical vectors, but was unable to take the mean of the factor and character vectors. In the case of the logical vector, R converted "TRUE" and "FALSE" values to 0s and 1s before calculating a mean. 



#### Converting data frame vectors to numeric
Convert the logical, character, and factor vectors in the data set to numeric vectors.  
```{r converting_to_numeric, eval = FALSE}
as.numeric(pull(prob1_df, samp_gt_0))
as.numeric(pull(prob1_df, char_vec))
as.numeric(pull(prob1_df, factor_vec))
```

R is able to convert the logical and factor vectors to numeric vectors, but was unable to do so for the character vector. As discussed earlier, R converted the logical vector to a numeric vector by changing "TRUE" and "FALSE" values to 0s and 1s. In the case of the factor vector, R assigned different numeric values (i.e. 1, 2, or 3) to the different levels of this variable. 

Therefore, in order to perform a mathematical operation on a factor vector (ex. calculating a mean), one would first need to convert the factor vector to a numeric vector. In the case of a logical vector, R will automatically do the appropriate conversion and therefore no additional function (as.numeric) needs to be applied before calculating a mean. 




#### Converting data frame vectors to logical and multiplying
Convert the the logical vector to numeric, and multiply the random sample by the result. Then,
convert the logical vector to a factor, and multiply the random sample by the result.
Lastly, convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result.

```{r convert_and_multiply}
as.numeric(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
as.factor(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
as.numeric(as.factor(pull(prob1_df, samp_gt_0))) * pull(prob1_df, samp)

```


## Problem 2 

#### Loading the data set
Load the penguins data set into R.
```{r load_penguins}
data("penguins", package = "palmerpenguins")
```



#### Describing the data set
Familiarize yourself with the penguins data set (by using ?penguins). Describe the data set.

The penguins data set includes information (collected between 2007-2009) about size measurements for adult (male and female) foraging penguins near Palmer Station, Antarctica (at any of the following islands: `r levels(pull(penguins, island))`). The data set contains `r ncol(penguins)` columns, `r nrow(penguins)` rows, and consists of the following variables: `r names(penguins)`. 

The mean flipper length for penguins in this data set is `r round(mean(pull(penguins, flipper_length_mm), na.rm = T), digits = 2)` mm  with a standard deviation of `r round(sd(pull(penguins, flipper_length_mm), na.rm = T), digits = 2)` mm. 

The mean bill length of penguins in this data set is `r round(mean(pull(penguins, bill_length_mm), na.rm = T), digits = 2)` mm with a standard deviation of `r round(sd(pull(penguins, bill_length_mm), na.rm = T), digits = 2)` mm.

The mean bill depth of penguins in this data set is `r round(mean(pull(penguins, bill_depth_mm), na.rm = T), digits = 2)` mm with a standard deviation of `r round(sd(pull(penguins, bill_depth_mm), na.rm = T), digits = 2)` mm.

The mean body mass of penguins in this data set is `r round(mean(pull(penguins, body_mass_g), na.rm = T), digits= 2)` g with a standard deviation of `r round(sd(pull(penguins, body_mass_g), na.rm = T), digits = 2)` g.




#### Creating a scatterplot of the data set
Create a scatterplot of the penguins data set with appropriate title, axes, and color differentiation between species. 

```{r create_scatterplot}

penguins %>% 
    ggplot(aes(bill_length_mm, flipper_length_mm, color = species)) + 
    geom_point() +
    ggtitle("Penguins Scatterplot") +
    xlab("Bill Length (mm)") +
    ylab("Flipper Length (mm)") + 
    theme(plot.title = element_text(hjust = 0.5))

```



#### Exporting the scatterplot
Export scatterplot as a pdf using ggsave.
```{r export_scatterplot }
ggsave("hw1_plot.pdf")
```






