---
title: "Homework 1"
author: Yaa Ababio
output: github_document
---

This is my solution to HW1.  

```{r setup, include= false}
library(tidyverse)
```


## Problem 1

Create a data frame comprised of:
-a random sample of size 10 from a standard Normal distribution
-a logical vector indicating whether elements of the sample are greater than 0
-a character vector of length 10
-a factor vector of length 10, with 3 different factor “levels”.

```{r}
prob1_df = 
  tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vec = factor(c("small", "small", "small", "med", "med", "med", "med", "large", "large", "large"))
  )
```

Take the mean of each variable in my data frame.
```{r}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```

I was able to take the means of the numeric and logical vectors, but was unable to take the mean of the factor and character vectors. I was able to take the mean of the logical variable because R converted "TRUE" and "FALSE" values to 0s and 1s before calculating a mean. 

```{r}
as.numeric(pull(prob1_df, samp))
as.numeric(pull(prob1_df, samp_gt_0))
as.numeric(pull(prob1_df, char_vec))
as.numeric(pull(prob1_df, factor_vec))
```


## Problem 2 